---
title: "Gov 50 Final Project"
author: Allison Jones 
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---

## Milestone 1 (Project thoughts) 

I am interested in exploring data related to types of Halloween candy and their popularity among consumers. In particular, I am curious to learn if Halloween candy containing nuts tends to be more popular than nut-free candy. The study’s findings might help consumers make informed choices when purchasing Halloween candy by ensuring consumers choose treats that are more likely to be enjoyed by trick-or-treaters.

## Milestone 2 (Project proposal)

Does the presence of peanuts or nuts in Halloween candy affect its popularity in matchups? I hypothesize that Halloween candies containing peanuts or nuts will be more popular in matchups compared to candies without peanuts or nuts. This hypothesis is rooted in the belief that the presence of peanuts or nuts adds a desirable nutty flavor and crunchy texture to candies that may contribute to an appealing sensory experience, which leads to an increase in candy popularity. My sample is comprised of 86 different Halloween candies with each candy being the unit of analysis. The explanatory variable of interest is whether the candy contains peanuts or nuts, and it is coded =1 for candies that contain peanuts or nuts and =0 for candies that do not contain peanuts or nuts. My outcome variable of interest is the popularity of each Halloween candy, as measured by the percentage of matchups won by each candy in an online preference matchup. This variable will indicate how well each candy performed in terms of winning matchups against other candies. For instance, if the variable was coded  = 67, it would mean that the candy won 67% of the matchups it was involved in. This variable is measured from Walt Hickey’s data set on Halloween candy preferences and relies on participant responses collected from the online preference matchup that included 8,371 different IP addresses that participated in voting on approximately 269,000 randomly generated matchups. In these matchups, participants were presented with pairs of candies and asked to choose their preferred candy from each pair. The popularity of each candy was determined by calculating the percentage of matchups it won. If I observe a significantly higher percentage of matchups won by candies containing peanuts or nuts compared to candies without peanuts or nuts, this would provide support for my hypothesis. Conversely, if I observe that candies without peanuts or nuts consistently win a significantly higher percentage of matchups or if there is no significant difference in the popularity of candies with and without peanuts or nuts, this would provide evidence against my hypothesis. When I run my regression of the count of the popularity of Halloween candies in matchups, a positive and significant coefficient for the presence of peanuts or nuts in the candy would suggest support for my hypothesis.

## Milestone 3 (Data Visualization)
```{r echo = FALSE}
library(tidyverse)
library(ggplot2)
library(dplyr) 

candy <- read.csv("candy-data.csv")
```

```{r, echo = FALSE, fig.width = 8, fig.height = 11}

candy_rankings <- candy |> 
  mutate(`Types of Candy` = if_else(peanutyalmondy == 1, "Nut Candy", "Nut-Free Candy")) |>
  filter(! competitorname %in% c("One dime", "One quarter"))

candy_rankgings_graph <- candy_rankings |> 
  arrange(desc(winpercent)) |> 
  slice_max(winpercent, n = 25) |>
  ggplot(mapping = aes(x = winpercent, y = fct_reorder(competitorname, winpercent), color = `Types of Candy`, fill =`Types of Candy`)) + 
  geom_col() + 
  scale_fill_manual(values = c("steelblue1","indianred1")) +
  labs(x = "Percentage of Matchups Won by Candies", y = "Name of Candies", title = "Percentage of Matchups Won by Nut Candies vs. Nut-Free Candies", caption = "Walt Hickey's 538 Data Set")
candy_rankgings_graph
```

